---
title: "ggplot title"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

```
  
It is possible to add a title to a ggplot2 chart using the ggtitle() function. It works as follow:
  
```{r Method 1: the Forecats library, include=TRUE}
# library & data
library(ggplot2)
data <- data.frame(x=rnorm(100))
 
# Basic plot with title
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("A blue Histogram") +
  theme_minimal()
```
  
## Title on several lines
It is a common need to set the title on several lines. To add a break and skip to a second line, just add a \n in the text.
```{r, include=TRUE}
# title on several lines
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("A blue \nHistogram") +
  theme_minimal()
```
  
## Title appearance and position
Here are 2 tricks to control text appearance and its position. Both features are controled thanks to the plot.title argument of the theme() function. Appearance can be controlled with option such as family, size or color, when position is controlled with hjust and vjust.  
```{r, include=TRUE}
# Custom title appearance
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("A blue Histogram") +
  theme_minimal() +
  theme(
    plot.title=element_text(family='', face='bold', colour='purple', size=26)
  )
# Adjust the position of title
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.8) +
  ggtitle("With title on the right") +
  theme_minimal() +
  theme(
    plot.title=element_text( hjust=1, vjust=0.5, face='bold')
  )
```
  
## Customize a specific word only
  
If you want to highlight a specific set of words in the title, it is doable thanks to the expression() function.

```{r, include=TRUE}
# Custom a few word of the title only:
my_title <- expression(paste("A ", italic(bold("blue")), " histogram"))
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.8) +
  ggtitle(my_title) +
  theme_minimal()
```

